Optimal (Multiway) Spatial Joins

### 2024-05-30 msj.tex
yufei

    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
}
See Figure~\ref{fig:guard} for an illustration.

@@

\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
\myeqn{
    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
}
We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$. When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
\myeqn{
    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \label{eqn:hv:type1:topsliced}
}
and call it the {\em top-sliced rectangle} of $r'$.

@@

Next, we will prove 
\myeqn{
    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out. \label{eqn:hv:type1:properties:proof-help1}
} 
which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$. Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}). This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}. By statement (2), the $\J^*$ computed must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$. This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}

>>>>

    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
}
See Figure~\ref{fig:guard} for an illustration. It is worth mentioning that since a horizontal segment $h$ is a degenerated rectangle, notations such as $\xleft(h)$ and $\xright(h)$ are well-defined.

@@

\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
\myeqn{
    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
    \label{eqn:hv:type1:maxtop}
}
We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$. When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
\myeqn{
    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \label{eqn:hv:type1:topsliced}
}
and call it the {\em top-sliced rectangle} of $r'$.

@@

Next, we will prove 
\myeqn{
    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out. \label{eqn:hv:type1:properties:proof-help1}
} 
which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$. Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}). This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}. By statement (2), the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$. This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} \\
\underline{\em Proof of Statement (1).} We will map each $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t}) = \cross_H(r^*)$. After this, we can then derive $\sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out$, where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.

The mapping is designed as follows. Consider an arbitrary $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$. Recall that $r^*$ is the top-sliced of some rectangle $r'$. Specifically, if $r' = [x_1,x_2]\times [y_1,y_2]$, then $r^* = [x_1,x_2]\times [y_1,\maxtop(r')]$. By the definition of $\maxtop(r')$ in \eqref{eqn:hv:type1:topsliced}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_{\bm{t}})$. We map $r^*$ to $\bm{t}$.

%As $\gbot(\bm{t}) = r'$, the rectangle $B_{\bm{t}}$ can be represented as $[\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1, \maxtop(r')]$.

Next, we will prove that a segment $h \in \cross_H(r^*)$ if and only if $h \in \dcross_H(\bm{t})$. Let us start with the ``only-if direction'': assuming $h \in \cross_H(r^*)$, we aim to show $h \in \dcross_H(\bm{t})$. For this purpose, we need to show that $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
\myitems {
    \item First, we explain why $h$ crosses $r'$. Because $r'$ covers $r^*$, by the fact $h$ crossing $r^*$, we know that $h$ must intersect with $r'$. On the other hand, as $h$ crosses $r^*$, it must hold that $\xleft(h) < x_1 < x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$). As $[x_1, x_2]$ is also the x-range of $r'$, using the fact of $h \cap r' \ne \emptyset$, we can assert that $h$ crosses $r'$.

    \item Then, we explain why $h$ crosses $B_{\bm{t}}$. Note that because $h$ crosses $r'$ (which we just proved) and $B_{\bm{t}}\subseteq r'$, we know that $B_{\bm{t}}$ cannot contain any endpoint of $h$. Let $y_h$ be the y-coordinate of $h$, and take the point $p = (\xleft(B_{\bm{t}}), y_h)$. We will prove $p \in h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$ and, hence, $h$ must cross $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$, as mentioned).

    Since $\gbot(\bm{t}) = r'$, the bottom segment of $B_{\bm{t}}$ must be contained in $r'$. Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$. As proved earlier, $\xleft(h) < x_1 < x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting $p \in h$. To prove $p \in B_{\bm{t}}$, it remains to show $y_h \in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$. Since $h$ intersects $r^*$, $y \in [y_1 , \maxtop(r')]$ --- recall that $[y_1, \maxtop(r')]$ is the y-range of $r^*$ --- where $\maxtop(r') = \ytop(B_{\bm{t}})$. Since $\gbot(\bm{t}) = r'$, $y_1 = \ybot(B_{\bm{t}})$. Hence, $y\in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$, and $p\in B_{\bm{t}}$.

}

Conversely, let $h\in \dcross_H(\bm{t})$, which means that $h$ crosses $r'$ and $B_{\bm{t}}$. Thus, $r'$ does not contain any endpoints of $h$. Since $r^*\subseteq r'$, $r^*$ does not contain any endpoints of $h$ neither.
Next, we prove that $B_{\bm{t}} \subseteq r^*$. Since $h \cap B_{\bm{t}}\neq \emptyset$, this statement leads to $h \cap r^* \neq \emptyset$ and $h$ crosses $r^*$. Recall that $r^* = [x_1,x_2], [y_1, \maxtop(r')]$, and $B_{\bm{t}} = [\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1,\maxtop(r')]$. Since $B_{\bm{t}}$ is a subset of $r'$, we have $\xleft(B_{\bm{t}}) \subseteq [x_1,x_2]$. Therefore, $B_{\bm{t}} \subseteq r^*$, and $h\in \cross_H(r^*)$.

It remains to show that we will not map two distinct rectangle $r_1^*, r_2^*\in \bigcup_{i = 1}^{k-2}R_i^*$ to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$. Suppose to the contrary that $r_1^*, r_2^*$ are mapped to the same tuple $\bm{t}$. Let $r_1^*, r_2^*$ be the top-sliced rectangles of $r_1'$ and $r_2'$ respectively. Since $r_1^*$ and $r_2^*$ are disinct, $r_1'$ is not the same as $r_2'$. By our mapping, $\gbot(B_{\bm{t}})$ is $r_1'$ and $r_2'$. Since $\gbot(B_{\bm{t}})$ is unique for each tuple $\bm{t}$, we have a contradiction. Therefore, each rectangle $r^*\in \bigcup_{i = 1}^{k-2}R_i^*$ is mapped to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
